A practical look at genetic algorithms

### 2015-11-26 evolutionaryComputation/practicals/report.tex
added graphs section
did some graph exporting
created proper latex tables and now also ascii graphs
svg graphics now automaticly created by the code
fixed sentence
added extra text, removed a bunch of things I'm no longer uncertain about and added a begin to the uml diagram which I can make in latex aperantly.

\documentclass{article}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

@@

\[B(x_1\dots x_k)=\left\{
	\begin{matrix}
		k & \mbox{ if } &   CO(x_1\dots x_k) = k \\
		k - d - \frac{k-d}{k-1} CO(x) & \mbox{ if } & CO(x_1\dots x_k) < k \\
	\end{matrix}
\right. \]
\section{Methodology}
\subsection{Build tool}
I use Gradle for building the project. Its the most advanced build platform
available and makes it easy to add dependencies to the project.
For example I use slf4j for logging.

Its also handy in that it comes with a self installing script that can be run
from any platform, Gradlew. So running the code I wrote on a different machine
becomes a trivial task, Gradle makes sure the right dependencies are fetched.

\subsubsection{Run guide}
As said before its a trivial task to run the code. Downloading and installing
of dependencies and even gradle happens automatically with help of gradlew.
So the first run is slower, depending on your internet speed.

\subsubsection{Libraries used}

\subsection{Language choice}
I chose to use Scala to implement this project. The main reason being that I've
played with Scala before but never have done anything constructive with it.

@@

Other options I considered were:
\begin{itemize}
	\item Java, which I'm very familiar with but I find quite boring to use.
		Its very verbose.
	\item C\#, the language is good but setting it up is difficult since I
		use Linux, it also has no Gradle support which makes it clunky,
		you'd have to rely on .net libraries only which is limited. It
		also prefers object orientated  rather than functional programming,
		Scala prefers functional over object orientated.
		Scala's syntax is also more concise.
	\item Rust/Go, I'm curious about both these languages, but I'm not at
		all familiar with them, so I probably should do first a little
		hobby project with them before starting doing something serious like
		this, I've already done this with Scala before.
\end{itemize}
\subsection{Architecture}
The architecture was created incrementally, I just implemented the functions
without thinking much about how to combine them.  I first started with the
fittest functions. Then I went on with selection, and finally I did the
crossover.
I just implemented one or 2 options of these and then created an evolution
class that accepted functions with the right prototype. So now you can
create an evolution by giving it functions.

@@

\subsection{Uncertainties}
Should the elitism selection use tournement selection?
When precisly do we find the optimal solution?
	Should I wait for convergence or just say a string
	of all ones is good enough?
	Is 1 solution good enough or should all members have
	that solution
	I'm assuming 1 solution and that I can give a string
	of just ones.
How many generations to consider. I chose arbitrarly 10.

About randomly linked:
	There is another way to do it: for every member randomly link
	differently, and randomly pick 1 randomly linked instructionset
	for offspring generation (or do the strongest).
	Currently I do the global randomly linked, every run will however
	have the same randomly linked value, I'm aware of this.

I'm not quite sure about the graphs. I have to show the number of successes
out of the 30 runs, but this is already set in stone, because its either
29 or 30, or its no successes.

@@

However once I implemented this, I came to the conclusion it was not faster.
The overhead for delegating and creating threads was more than the parilization
gained.

@@

\subsubsection{Evaluate each member on creation}
Members are going to be evaluated always, initially I would evaluate the member
everytime it needed to be evaluated, but in later version I just aded the
evaluation to the case class. Its not going to change anyways.

@@

To reduce the amount of string comparisons, first the hascode function is called
This will cache itself, I know this because I inspected the source code of the
String class in Java.
\section{Execution}
%TODO:
% 1. Experiments
% Use stoastic within the algoritm
%	Draw a graph from the experiments
%		* pop size on x, # succes on Y
% maybe some more testing of the crossovers?
% also try to implement a threading scheme??

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz-uml}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{svg}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

@@

\[B(x_1\dots x_k)=\left\{
	\begin{matrix}
		k & \mbox{ if } &   CO(x_1\dots x_k) = k \\
		k - d - \frac{k-d}{k-1} CO(x) & \mbox{ if } & CO(x_1\dots x_k) < k \\
	\end{matrix}
\right. \]
\section{Methodology}
\subsection{Language choice}
I chose to use Scala to implement this project. The main reason being that I've
played with Scala before but never have done anything constructive with it.

@@

Other options I considered were:
\begin{itemize}
	\item Java, which I'm very familiar with but I find quite boring to use.
		Its very verbose.
	\item C\#, the language is good but setting it up is difficult since I
		use Linux, it also has no Gradle support which makes it clunky,
		you'd have to rely on .net libraries only which is limited. It
		also prefers object orientated  rather than functional programming,
		Scala prefers functional over object orientated.
		Scala's syntax is also more concise.
	\item Rust/Go, I'm curious about both these languages, but I'm not at
		all familiar with them, so I probably should do first a little
		hobby project with them before starting doing something serious like
		this, I've already done this with Scala before.
\end{itemize}

\subsection{Architecture}
The architecture was created incrementally, I just implemented the functions
without thinking much about how to combine them.  I first started with the
fittest functions. Then I went on with selection, and finally I did the
crossover.

\begin{tikzpicture}
\begin{umlpackage}[x=0,y=0]{ec}
	\umlemptyclass[x=3]{Member}
\umlemptyclass{Population}
\umlemptyclass[x=6]{PairedPopulation}
%\umluniassoc[geometry=-|,arg1=x,mult1=1,pos1=1.9,arg2=y,mult2=*,pos2=0.2]{Population}{Member}
\end{umlpackage}
\end{tikzpicture}

I just implemented one or 2 options of these and then created an evolution
class that accepted functions with the right prototype. So now you can
create an evolution by giving it functions.

Most of the GA architecture was already in place before the assingment became
available. I did this based on the notes I took in class. When the assignment
became available I was almost completly done with the GA. There were some bugs
in my code however and I interperted some things wrong. I also didn't write
anything for an experiment. So I still needed to do that.

@@

\subsubsection{Build tool}
I use Gradle for building the project. Its the most advanced build platform
available and makes it easy to add dependencies to the project.
For example I use slf4j for logging.

Its also handy in that it comes with a self installing script that can be run
from any platform, Gradlew. So running the code I wrote on a different machine
becomes easy, Gradle makes sure the right dependencies are fetched.

\paragraph{Running it}

Open a shell and move to the project folder then type:

\lstset{language=Bash}
\begin{lstlisting}[frame=single]
	./gradlew run
\end{lstlisting}

\subsubsection{Libraries used}
In here I will describe why I chose to use these libraries
The details of the libraries can be found in the gradle file (which
version and what maven group):

\begin{lstlisting}[frame=single]
	cat code/build.gradle
\end{lstlisting}

\begin{itemize}
	\item The scala standart library, because I have to when using scala.
	\item slf4j, a simple logging facade, this allows runtime switching of
		logging schemes, for example if you want to log into a database or
		a file instead of stdout you can do this relativly easy with slf4j
		This is not necisary right now but the api is simple enough to use
		everywhere.
	\item scalaplot. A graph plotting library for scala. I didn't want to
		do this myself, and this library supports svg plotting and ascii
		plotting.
	\item scalatest. A testing library for scala, it has some pretty advanced
		features but I use it mostly like JUnit.
\end{itemize}
\subsection{Uncertainties}
About randomly linked:
	There is another way to do it: for every member randomly link
	differently, and randomly pick 1 randomly linked instructionset
	for offspring generation (or do the strongest).
	Currently I do the global randomly linked, every run will however
	have the same randomly linked value, I'm aware of this.

@@

However once I implemented this, I came to the conclusion it was not faster.
The overhead for delegating and creating threads was more than the parilization
gained.  After 10 minutes of waiting I decided to drop this idea 
since without parilization it took about 2 minutes (I tried this
on just one run, so it shouldn't take that long).

@@

So now all expiriments are run at the same time. This probably increases the
average time of each 'run' somewhat (not that much because the concurent
collection does \href{http://docs.scala-lang.org/overviews/parallel-collections/configuration.html}{load balancing})
It alo drasticly decreases the amount of time I have to wait. Which is quite
important, especially in the development phase. (although I also used
other tricks, such as instead of 30, just doing 3 runs)

The paralization of experiments can be removed by deleting `par' on line 64
of Main.scala

\subsubsection{Evaluate each member on creation}
Initially my members just contained the string with the `genes'. This I passed
around and would eveulate when needed. However I realized that I could do this
evaluation eagerly on creation, to reduce the evaluation calls. A member has
to be eveluated anyways, so doing this on creation and just putting in the
evaluation saves some heavy evaluation operations.

@@

To reduce the amount of string comparisons, first the hascode function is called
This will cache itself, I know this because I inspected the source code of the
String class in Java.

I was very proud for coming up with  this idea, but when I ran the expirment with
and without hascode I came to the conclusion that first doing a hascode operation
significantly *slows* execution. This becomes worse as population size increases.

\section{Execution}
So I can execute in parralel, I can easily add whatever function I want, so
I decided to just run a whole bunch of experiments and see what comes out.

\input{result.tex}
